Numerical Investigation of Viscous Flow Instabilities in Multiphase Heterogeneous Porous Media

### 2018-07-17 Papers/ViscousInstabilities/Manuscript_ViscousInstabilities_KC.tex
Last Revision: small changes to improve readibility

In this work, all test-cases were performed in idealised geometries discretised with unstructured triangular and tetrahedral mesh using the \PN[1]{2}, \PN[1]{1} (Fig.~\ref{fig:fem_elem}, Sections~\ref{section:results_initial_model_validation}-\ref{section:results_hete_fix_adapt}) and \PNDG[1]{1} (Section~\ref{section:results_3D})  FE-pairs. An implicit Crank-Nicolson time-stepping scheme was used with {\it a posteriori} adaptive time-step size targeting a maximum Courant-Friedrichs-Lewy condition \citep[CFL,][]{courant_1941} of 2. In most simulations, the domain was initially fully saturated with non-wetting fluid which was displaced by a pure (wetting) fluid at a prescribed initial velocity $\left(u^{\circ}\right)$. For simplicity, the porosity ($\phi$) of the domain was kept constant at 0.2 in all simulations, whereas the absolute permeability ($\mathbf{K}$) varied in space, \ie $\mathbf{K}=\mathbf{K}\left(x_{i}\right)$. Fluids are assumed incompressible, also gravity and capillary pressure were neglected. Initial set-up for the numerical simulations is summarised in Table \ref{table:setup}.

>>>>

In this work, all test-cases were performed in idealised geometries discretised with unstructured triangular and tetrahedral mesh using the \PN[1]{2}, \PN[1]{1} (Fig.~\ref{fig:fem_elem}, Sections~\ref{section:results_initial_model_validation}-\ref{section:results_hete_fix_adapt}) and \PNDG[1]{1} (Section~\ref{section:results_3D})  FE-pairs. An implicit Crank-Nicolson time-stepping scheme was used with {\it a posteriori} adaptive time-step size targeting a maximum Courant-Friedrichs-Lewy condition \citep[CFL,][]{courant_1941} of 2. In most simulations, the domain was initially fully saturated with non-wetting fluid which was displaced by a pure (wetting) fluid at a prescribed initial velocity $\left(\mathbf{u}_{\alpha}^{\circ}\right)$. For simplicity, the porosity ($\phi$) of the domain was kept constant at 0.2 in all simulations, whereas the absolute permeability ($\mathbf{K}$) varied in space, \ie $\mathbf{K}=\mathbf{K}\left(x_{i}\right)$. Fluids are assumed incompressible, also gravity and capillary pressure were neglected. Initial set-up for the numerical simulations is summarised in Table \ref{table:setup}.

### 2018-07-17 Papers/ViscousInstabilities/article_figure1.tex
Last Revision: small changes to improve readibility

%%%%
%%%%  FIGURE 
%%%%
\begin{figure}[h]
\vbox{\vspace{0.cm}
\hbox{\hspace{2cm}
\includegraphics[width=.5\textwidth]{./Pics/P1DGP2.pdf}}
\vspace{0.cm}
\hbox{\hspace{5cm}(a)}
\vspace{0.5cm}
\hbox{\hspace{1cm}
\includegraphics[width=.75\textwidth]{./Pics/element_n_2b.pdf}}
\vspace{0.cm}
\hbox{\hspace{3.5cm}(b)\hspace{3cm}(c)}}
%\vbox{\includegraphics[width=.5\textwidth]{./Pics/P1DGP2.pdf}}
\caption{2D representation of \PN[n]{m} element-pairs used in this work: (a) set of \PN[1]{2} FE with shaded areas as control volumes across two contiguous elements (blue and white circles represent pressure and velocity nodes, respectively); Detailed (b) \PN[1]{2} and (c) \PN[1]{1} element-pairs.} 
\label{fig:fem_elem}
\end{figure}
\clearpage

>>>>

%%%%
%%%%  FIGURE 
%%%%
\begin{figure}[h]
\vbox{\vspace{0.cm}
\hbox{\hspace{2cm}
\includegraphics[width=.5\textwidth]{./Pics/P1DGP2.pdf}}
\vspace{0.cm}
\hbox{\hspace{5cm}(a)}
\vspace{0.5cm}
\hbox{\hspace{1cm}
\includegraphics[width=.75\textwidth]{./Pics/element_n_2b.pdf}}
\vspace{0.cm}
\hbox{\hspace{3.5cm}(b)\hspace{3cm}(c)}}
%\vbox{\includegraphics[width=.5\textwidth]{./Pics/P1DGP2.pdf}}
\caption{2D representation of \PN[n]{m} element-pairs used in this work: (a) set of \PN[1]{2} FE with shaded areas as control volumes across two contiguous elements (blue and white circles represent pressure and velocity nodes, respectively). Detailed (b) \PN[1]{2} and (c) \PN[1]{1} element-pairs.} 
\label{fig:fem_elem}
\end{figure}
\clearpage
